SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2018-07-25 scipy-1.0/ckdtree.tex
Fuse subsections for kdtree tech improvements; add a few refs.e

The cKDTree module is rewritten in C++ with templated classes.

The support to a box of periodic boundary condition is added. A periodic
condition is typically used in computer simulations of physical processes.

\subsubsection{k-NN}

@@

\subsubsection{sparse distance matrix}

In 2015 SciPy added support for generating approximated sparse distance matrices 
between \texttt{KDTree} objects (sets of points organized into a binary space 
partitioning data structure\cite{Bentley:1975:MBS:361002.361007}). 
The user provides a maximum distance value
above which all distances between points in two provided \texttt{KDTree} objects
are set to 0. The distance metric used between k-d trees is not constrained
to the conventional L2 (Euclidean) norm---any Minkowski p-norm value
bounded by 1 and infinity is valid. By default, a dictionary of keys
based sparse matrix is the returned data structure. Previously-published
results had suggested that a hashing approach to sparse matrix assembly
is 7 times faster than constructing with compressed row format (CSR)\cite{10.1007/978-3-540-75755-9_107}.
The dictionary object maps row and column indices to elements of the sparse matrix, which is
efficient for matrix construction because all zero values are skipped in
the dictionary mapping. The C++ level sparse matrix construction releases the Python
global interpreter lock for increased performance. Once constructed, the
dictionary of keys sparse matrix has the amortized constant time complexity 
($O(1)$) for distance value retrieval that one would expect from a 
dictionary\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic
operations on the sparse matrix, SciPy allows the dictionary of keys
sparse matrix to be directly converted to the common CSR, CSC, and COO
data structures.

\subsubsection{Pair-Counting}

The cKDTree module implements the dual tree counting algorithm in (cite).

An improvement to the pair counting algorithm is added to improve the scale
with the number of bins.

\textbf{Add a Figure to show the scaling, before and after.}

The cKDTree can now be augmented by weights. Weighted paircount is essential
in many scientific applications, e.g. computing correlation function of galaxies.

\textbf{perhaps give an example or some formula.}

(cite / mention faster implementions of paircounting algorithms / treecorr, corrfunc)

>>>>

The cKDTree module was rewritten in C++ with templated classes, and support for
periodic boundary conditions was added. A periodic boundary condition is typically 
used in computer simulations of physical processes.

@@

Also in 2015, SciPy added support for generating approximated sparse distance matrices 
between \texttt{KDTree} objects (sets of points organized into a binary space 
partitioning data structure\cite{Bentley:1975:MBS:361002.361007}). 
The user provides a maximum distance value
above which all distances between points in two provided \texttt{KDTree} objects
are set to 0. The distance metric used between k-d trees is not constrained
to the conventional L2 (Euclidean) norm---any Minkowski p-norm value
bounded by 1 and infinity is valid. By default, a dictionary of keys
based sparse matrix is the returned data structure. Previously-published
results had suggested that a hashing approach to sparse matrix assembly
is 7 times faster than constructing with compressed row format (CSR)\cite{10.1007/978-3-540-75755-9_107}.
The dictionary object maps row and column indices to elements of the sparse matrix, which is
efficient for matrix construction because all zero values are skipped in
the dictionary mapping. The C++ level sparse matrix construction releases the Python
global interpreter lock for increased performance. Once constructed, the
dictionary of keys sparse matrix has the amortized constant time complexity 
($O(1)$) for distance value retrieval that one would expect from a 
dictionary\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic
operations on the sparse matrix, SciPy allows the dictionary of keys
sparse matrix to be directly converted to the common CSR, CSC, and COO
data structures.

The cKDTree module implements a dual tree counting algorithm\cite{pmlr-v28-curtin13},
with an improvement to the pair counting algorithm to improve the scaling
with the number of bins. The cKDTree can now be augmented by weights, with 
weighted paircount essential in many scientific applications, e.g. computing 
correlation function of galaxies\cite{0004-637X-750-1-38}.

\textbf{Add a Figure to show the scaling, before and after.}
\textbf{perhaps give an example or some formula.}
(cite / mention faster implementions of paircounting algorithms / treecorr, corrfunc)
